Multi-domain, multilingual translation with latent multi-task group dropout

### 2022-01-13 2022/acl/acl_latex.tex
add answers for todos, local changes

\subsection{System design and configuration}
\subsubsection{Multidomain translation systems}
Our systems for the multidomain experiments are designed as follows:
\begin{itemize}
\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads; 6 layers in the encoder; 6 layers in the decoder.\fyDone{How many layers}
\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
\item Transformer using Latent multi-task group dropout (\system{LaMGD} Transformers):\fyDone{Explain acronym} There is no change in the architecture. We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 out of the 16 groups are selected. The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
\end{itemize}

We set the dropout probability to 0.1. We train the multi-domain Transformer model for 20k iterations with a batch size of 12k tokens. We train \system{LaMGD} Tranformer for 300k iterations with a batch size of 12k tokens: these extra iterations are justified by the need for the \system{LaMGD} Tranformer to learn both the sub-networks and the parameters of the underlying Transformer.\fyDone{Have we converged?} Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline.

@@

We set the dropout probability to 0.3. We train the multilingual Transformer model for 40k iterations with a batch size of 9600 tokens on 16 V100 GPUs as in \citet{Gong21adaptive}. We train \system{LaMGD} Tranformer for 50k iterations with the same batch size. Finally, we finetune the language-specific Adapters for 5k iterations.

@@

The temperature parameter $\tau$ for the $\operatorname{Soft-Top-K}$ operator is gradually decreased from $0.5$ to $0.2$ according to the following policy:
\begin{align*}
\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \},
\end{align*}
in which $r=0.0001$.

Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments. The choice of the values for our hyper-parameters was not optimized and selected by suggestions from previous studies \citep{Jang17categorical,Gong21pay,Gong21adaptive}.

For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \citep{Post18call}.\fyDone{Check this}

\subsection{Datasets}
\subsubsection{Multidomain translation}
We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{The raw data can be downloaded from \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
    \cline{2-7} 
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
  }
\label{tab:Corpora-chap4}
\end{table*}
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks borrowing the multilingual translation datasets from past studies. More precisely, we used: 
\begin{itemize}
\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations from \citet{qi18when} between English and eight \emph{related languages}.
\item TED8-Diverse. The dataset consists of parallel sentences between English and eight \emph{diverse languages} as in \citet{Wang20balancing}.
\end{itemize}

@@

\subsection{Multidomain translation}
For these experiments, our main results are in Table ~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Tranformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters. Moreover, \system{LaMGD} Tranformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is carrefully finetuned and contains approximately 25M additional parameters per domain.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & 40.3 & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & 61.0 & 53.1 & 37.5 & 49.6 & 91.0 & 55.3 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & 40.3 & 60.4 & 52.4 & 39.0 & 52.4 & 87.5 & 55.3 \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}

@@

% \begin{figure*}[h!]
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include first image
%   \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
%   \caption{TED-Related}
%   \label{fig:phi-related}
% \end{subfigure}
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include second image
%   \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
%   \caption{TED-Diverse}
%   \label{fig:phi-diverse}
% \end{subfigure}
% \caption{$\Phi_L^{0}$ in O2M experiments}
% \label{fig:phi}
% \end{figure*}
\subsection{Ablation study}
We discuss here the choice of the hyper-parameters $k$, the number of activated nodes in each layer, and its impact on the sharing level between the tasks. Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.\Todo{add sharing results. }\Todo{check caption consistency}
\begin{table}[h!]
  \centering
  \begin{tabular}{|c|c|} \hline
    $k$ & \domain{avg}\\ \hline 
    8 & 18.1 \\
    10 & 19.15 \\
    12 & 19.33 \\
    14 & 19.44 \\
    \hline
  \end{tabular}
  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$. These results correspond to the multilingual setting (o2m-diverse).}
  \label{tab:K}
\end{table}

>>>>

\subsection{System design and configuration}
\subsubsection{Multidomain translation systems}
Our systems for the multidomain experiments are designed as follows:
\begin{itemize}
\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads; 6 layers in the encoder; 6 layers in the decoder.\fyDone{How many layers}
\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
\item Transformer using Latent multi-task group dropout (\system{LaMGD} Transformers):\fyDone{Explain acronym} There is no change in the architecture. We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 out of the 16 groups are selected. The number of parameters of the variational distribution is $L\times k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
\item \item Transformer using heuristic multi-task group dropout (\system{HMGD} Transformer): we share 320 nodes for every task, and reserve 32 nodes for each task (making $320 + 32*6 = 512$ nodes in total).
\end{itemize}

We set the dropout probability to 0.1. We train the multi-domain Transformer model for 200k iterations with a batch size of 12k tokens. The convergence of the standard Transformer is before 200K as its validation curve became flat near the 200K-th iteration. The \system{LaMGD} Tranformer converged after 300k iterations with a batch size of 12k tokens. The convergence of \system{LaMGD} is controlled by its validation curve. Finally, we plug domain adapters to the multidomain Transformer model and finetune them for 25k iterations using the same batch size as the baseline. With approximately the same amount of training iterations, \system{LaMGD} and Adapter-based Transformer are trained.

@@

We set the dropout probability to 0.3. We train the multilingual Transformer model for 40k iterations with a batch size of 9600 tokens on 16 V100 GPUs as in \citet{Gong21adaptive}. We train \system{LaMGD} Transformer for 50k iterations with the same batch size. The convergence of the models are controlled via their validation curves. Finally, we finetune the language-specific Adapters for 5k iterations.

@@

The temperature parameter $\tau$ for the $\operatorname{Soft-Top-K}$ operator is gradually decreased from $0.5$ to $0.2$ according to the following policy:
\begin{align*}
\tau = \operatorname{min}\{ 0.2, 0.5 * \exp^{-r*step} \},
\end{align*}
in which $r=0.0001$.
While \citet{Gong21pay,Gong21adaptive} fixed $\tau$ to be 0.2, we select an anneal policy for $\tau$ proposed by previous studies \citep{Jang17categorical}. Finally, we fix the weight of the entropy $\mathbb{H}(\text{softmax}(\Phi_l^d))$ to $0.0001$ in the training loss in every experiments

For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \citep{Post18call}.\fyDone{Check this}
\subsubsection{Latent variables' initialization}
We initialize the probabilistic distribution of the latent variables by uniform distribution. More precisely, we set $\Phi_l^d$, which generates the probability of the masks via the softmax activation function, to $0^{n_d}$.
\subsection{Datasets}
\subsubsection{Multidomain translation}
We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{The raw data can be downloaded from \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
    \cline{2-7} 
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
  }
\label{tab:Corpora-chap4}
\end{table*}
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks borrowing the multilingual translation datasets from past studies. More precisely, we used: 
\begin{itemize}
\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations from \citet{qi18when} between English and eight \emph{related languages}.
\item TED8-Diverse. The dataset consists of parallel sentences between English and eight \emph{diverse languages} as in \citet{Wang20balancing}.
\end{itemize}

@@

\subsection{Multidomain translation}
For these experiments, our main results are in Table ~\ref{tab:mdmt}, where we observe that the \system{LaMGD} Tranformer achieves a significant improvement (+2.78) over the generic \system{Transformer} system with zero extra parameters. Moreover, \system{LaMGD} Tranformer achieves performance that are equivalent on average to that of the \system{Adapter} sytems, which is carrefully finetuned and contains approximately 25M additional parameters per domain. Variational mask learned from data by \system{LaMGD} also outperforms heuristic dropout mask \system{HMGD} by 0.5 in average.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{7}{r|}} \hline
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[65m]} & \SB{40.3} & 59.5 & 49.8 & 36.4 & 49.0 & 80.0  & 52.5\\
    \revision{\system{Adapter}}   \hfill{\footnotesize[+151m]}  & 39.5 & \SB{61.0} & \SB{53.1} & 37.5 & 49.6 & \SB{91.0} & \SB{55.3} \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  & \SB{40.3} & 60.4 & 52.4 & \SB{39.0} & \SB{52.4} & 87.5 & \SB{55.3} \\ 
    \system{HMGD} Tranformer   \hfill{\footnotesize[+0m]}  & \SB{40.4} & 60.4 & 51.9 & 38.7 &	50.80 &	86.80 & 54.8 \\ 
    \hline
  \end{tabular}
  \caption{Multi-domain translation}
  \label{tab:mdmt}
\end{table*}

@@

% \begin{figure*}[h!]
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include first image
%   \includegraphics[width=0.9\textwidth]{phi_ted_related_o2m}  
%   \caption{TED-Related}
%   \label{fig:phi-related}
% \end{subfigure}
% \begin{subfigure}{0.5\textwidth}
%   \centering
%   % include second image
%   \includegraphics[width=0.85\textwidth]{phi_ted_diverse_o2m}  
%   \caption{TED-Diverse}
%   \label{fig:phi-diverse}
% \end{subfigure}
% \caption{$\Phi_L^{0}$ in O2M experiments}
% \label{fig:phi}
% \end{figure*}
\subsection{Ablation study}
We discuss here the choice of the hyper-parameters $k$, the number of activated nodes in each layer, and its impact on the sharing level between the tasks. Table ~\ref{tab:K} shows the variance of performance when the number of activated nodes is changed, and the sharing level between tasks decreases in consequence.\Todo{add sharing results. }\Todo{check caption consistency}
\begin{table}[h!]
  \centering
  \begin{tabular}{|c|c|c|} \hline
    $k$ & \domain{avg} & sharing rate \\ \hline 
    8 & 18.1 & 0.63\\
    10 & 19.15 & 0.73\\
    12 & 19.33 & 0.78\\
    14 & 19.44 & 0.88\\
    \hline
  \end{tabular}
  \caption{Variation of the performance w.r.t $k$, while we fix $n_p=16$. These results correspond to the multilingual setting (o2m-related).}
  \label{tab:K}
\end{table}
